{\bf Landslide: A New Race-Finding Tool for 15-410}

### 2016-02-10 landslide-lecture/2015/lecture.tex
Update lecture for S16

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2015, September 30}

@@

\begin{document}
\renewcommand{\inserttotalframenumber}{40}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

	\linegap
	When do we fear being preempted?
	\begin{itemize}
		\item Threads becoming runnable (\texttt{thr\_create()}, \texttt{cond\_signal()}, etc.)
			\begin{itemize}
				\item Preemptions may cause it to run before we're ready
			\end{itemize}
		\item Synchronization primitives (\texttt{mutex\_lock()}/\texttt{unlock()}, etc.)
			\begin{itemize}
				\item If buggy or used improperly\ldots
			\end{itemize}
		\item Unprotected shared memory accesses (``data races'')% Say: more on this later.
			\begin{itemize}
				\item May result in data structure corruption
			\end{itemize}
	\end{itemize}

@@

	Data races are {\em not necessarily} bugs, just highly suspicious!
	\begin{itemize}
		\item Bakery algorithm: Is {\tt number[i]=max(number[0],number[1])} bad?
		%\item Is unprotected {\tt foobar\_statistics\_counter++} bad?
		\item What about unprotected {\tt next\_thread\_id++}?
		\item ``If threads interleaved the wrong way here, it {\em might} crash later.''
		\begin{itemize}
			\item Hmmm...
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Iterative Deepening}
	Minimal state space includes only ``mandatory'' context switches
	\begin{itemize}
		\item e.g., {\tt yield()}, {\tt cond\_wait()}.
	\end{itemize}
	\vspace{0.29in}
	\begin{center}
		\includegraphics[width=0.35\textwidth]{tree0.pdf}
	\end{center}
\end{frame}

@@

	New tests
	\begin{itemize}
		\item {\tt broadcast\_test}
		\item {\tt paradise\_lost} (new this semester)
		\item {\tt mutex\_test} (new this semester)
	\end{itemize}
\end{frame}

@@

\begin{frame}{Last Semester}
	Spring 2015: 7 groups signed up to use Landslide; 5 found bugs
	%\begin{itemize}
	%	\item (that they didn't find during stress testing)
	%\end{itemize}
	%%%% hopefully!
	\linegap

\begin{frame}{A new treat since last semester}
	\texttt{mutex\_test} tries to make your mutexes fail or deadlock.
	\linegap

	Landslide will look for data races inside your mutex implementation.
	\begin{itemize}
		\item {\tt xchg}/{\tt cmpxchg}/{\tt xadd}
		\item {\tt mutex->whose\_turn\_is\_it = ...;}
	\end{itemize}
	\pause
	\linegap

	Tested 77 P2s from S'14, F'14, S'15
	\begin{itemize}
		\item 5 unrelated deterministic bugs (use-after-free, crash, etc)
			% TODO check these for false positives
			% But this was run after the arbiter fp deadlock avoidance
			% So I think it's ok
		\item 1 ``two threads got the lock at once'' race
			\begin{itemize}
				\item {\em ...that the TA who graded it didn't find!}
			\end{itemize}
		\item 7 deadlock races
		\item 4 other races (NULL crash, infinite loop, use-after-free)
	\end{itemize}
\end{frame}

@@

	To participate\dots
	\begin{itemize}
		\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item \url{http://tinyurl.com/landslide-p2-f15}
			\end{itemize}

@@

\subsection{Experimental Results}

\begin{frame}{More Evaluation}
	\textbf{Is dynamically adding ``data race'' PPs effective?}
	\begin{itemize}
		%\item 3 among preliminary bugs (9-15\%) required data races to expose.
		\item 6 among S'15 student bugs (46\%)
		\item 10 among {\tt mutex\_test} bugs (71\%)
	\end{itemize}
	\linegap

	{\bf Is ``iterative deepening'' better than state-of-the-art?}
	\begin{itemize}
		\item Control experiment: Just 1 state space, same CPU time
		\begin{itemize}
			\item PPs used: \texttt{mutex\_lock()}, \texttt{mutex\_unlock()}
			\item 110 minutes on 1 CPU
		\end{itemize}
		\item Of 33 total bug reports, control failed to find 10 (30\%).
		\begin{itemize}
			\item 3 required data race PPs to expose
			\item 1 ran out of time
			\item 6 obscured by different bug in same state space
		\end{itemize}
	\end{itemize}
\end{frame}

>>>>

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2016, February 10}

@@

\begin{document}
\renewcommand{\inserttotalframenumber}{39}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

	\linegap
	When do we fear being preempted?
	\begin{itemize}
		\item Threads becoming runnable (\texttt{thr\_create()}, \texttt{cond\_signal()}, etc.)
			\begin{itemize}
				\item Preemptions may cause it to run before we're ready
			\end{itemize}
		\item Synchronization primitives (\texttt{mutex\_lock()}/\texttt{unlock()}, etc.)
			\begin{itemize}
				\item If buggy or used improperly\ldots
			\end{itemize}
		\item Unprotected shared memory accesses (``data races'')% Say: more on this later.
			\begin{itemize}
				\item May result in data structure corruption
				\item More on this later...
			\end{itemize}
	\end{itemize}

@@

	Data races are {\em not necessarily} bugs, just highly suspicious!
	\begin{itemize}
		\item Bakery alg: Is {\tt number[i]=max(number[0],number[1])+1} bad?
		%\item Is unprotected {\tt foobar\_statistics\_counter++} bad?
		\item What about unprotected {\tt next\_thread\_id++}?
		\item ``If threads interleaved the wrong way here, it {\em might} crash later.''
		\begin{itemize}
			\item Hmmm...
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\begin{frame}{Iterative Deepening}
	Minimal state space includes only ``mandatory'' context switches
	\begin{itemize}
		\item e.g., {\tt yield()}, {\tt cond\_wait()}.
		% Say: When the test invokes yield, the kernel scheduler will pick
		% one thread to run next. With Landslide, we will pick ALL threads to
		% run next, in one parallel universe at a time.
	\end{itemize}
	\vspace{0.29in}
	\begin{center}
		\includegraphics[width=0.35\textwidth]{tree0.pdf}
	\end{center}
\end{frame}

@@

	New tests
	\begin{itemize}
		\item {\tt broadcast\_test}
		\item {\tt paradise\_lost} %(new this semester)
		\item {\tt mutex\_test} %(new this semester)
	\end{itemize}
\end{frame}

@@

\begin{frame}{Last Semester}
	%Spring 2015: 7 groups signed up to use Landslide; 5 found bugs
	%%\begin{itemize}
	%%	\item (that they didn't find during stress testing)
	%%\end{itemize}
	%%%%% hopefully!
	%\linegap

	%Among all groups, 38 total tests were run
	%\begin{itemize}
	%	\item 21 tests ran to completion without finding bugs
	%	\item 4 {\em deterministic} bugs
	%	\item 13 distinct {\em non-deterministic} bugs
	%		\begin{itemize}
	%			\item 6 in {\tt thr\_exit\_join} \small (use-after-free, deadlock, NULL, infinite loop...)
	%			\item 4 in {\tt paraguay} (deadlock, infinite loop)
	%			\item 2 in {\tt broadcast\_test} (deadlock, infinite loop)
	%			\item 1 in {\tt rwlock\_downgrade\_read\_test} (use-after-free)
	%		\end{itemize}
	%		%\begin{itemize}
	%		%	\item 6 of which required data-race PPs to expose
	%		%\end{itemize}
	%\end{itemize}
	Fall 2015: 22 groups signed up to use Landslide; 21 found bugs
	% 18 groups found races
	\begin{itemize}
		\item (that they didn't find during stress testing)
	\end{itemize}
	%%%% hopefully!
	\linegap

	Among all groups, 185 distinct tests were run
	\begin{itemize}
		\item 137 tests ran to completion without finding bugs
		\item 21 {\em deterministic} bugs
	\end{itemize}
	\pause % ask: "who can tell me why landslide might find a *deterministic* bug not found during past testing? A: use-after-free
	\linegap

	38 distinct {\em non-deterministic} bugs
	\begin{itemize}
			%\begin{itemize}
			%	\item 7 of which required data-race PPs to expose
			%\end{itemize}
		\item 11 groups (50\%) fixed all their bugs
			\begin{itemize}
				\item (as verified by running Landslide again -- not a guarantee!)
			\end{itemize}
		\item 2.7 bugs ``fixed'' (see above) on average per group
		\item Most ambitious group: 11 distinct races found + fixed!
	\end{itemize}
\end{frame}

%\begin{frame}{A new treat} %since last semester}
%	\texttt{mutex\_test} tries to make your mutexes fail or deadlock.
%	\linegap
%
%	Landslide will look for data races inside your mutex implementation.
%	\begin{itemize}
%		\item {\tt xchg}/{\tt cmpxchg}/{\tt xadd}
%		\item {\tt mutex->whose\_turn\_is\_it = ...;}
%	\end{itemize}
%	\pause
%	\linegap
%	
%	Last semester: 12 non-determinstic bugs among 8 groups
%	\linegap
%
%	%Tested 77 P2s from S'14, F'14, S'15
%	%\begin{itemize}
%	%	\item 5 unrelated deterministic bugs (use-after-free, crash, etc)
%	%		% TODO check these for false positives
%	%		% But this was run after the arbiter fp deadlock avoidance
%	%		% So I think it's ok
%	%	\item 7 deadlock races
%	%	\item 4 other races (NULL crash, infinite loop, use-after-free)
%	%	\item 1 ``two threads got the lock at once'' race
%	%		\begin{itemize}
%	%			\item {\em ...that the TA who graded it didn't find!}
%	%		\end{itemize}
%	%\end{itemize}
%\end{frame}

@@

	To participate\dots
	\begin{itemize}
		\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item \url{http://tinyurl.com/landslide-p2-s16}
			\end{itemize}

@@

%\subsection{Experimental Results}
%
%\begin{frame}{More Evaluation}
%	\textbf{Is dynamically adding ``data race'' PPs effective?}
%	\begin{itemize}
%		%\item 3 among preliminary bugs (9-15\%) required data races to expose.
%		\item 6 among S'15 student bugs (46\%)
%		\item 10 among {\tt mutex\_test} bugs (71\%)
%	\end{itemize}
%	\linegap
%
%	{\bf Is ``iterative deepening'' better than state-of-the-art?}
%	\begin{itemize}
%		\item Control experiment: Just 1 state space, same CPU time
%		\begin{itemize}
%			\item PPs used: \texttt{mutex\_lock()}, \texttt{mutex\_unlock()}
%			\item 110 minutes on 1 CPU
%		\end{itemize}
%		\item Of 33 total bug reports, control failed to find 10 (30\%).
%		\begin{itemize}
%			\item 3 required data race PPs to expose
%			\item 1 ran out of time
%			\item 6 obscured by different bug in same state space
%		\end{itemize}
%	\end{itemize}
%\end{frame}

\begin{frame}{Last-last Semester}
	Spring 2015: 7 groups signed up to use Landslide; 5 found bugs
	%\begin{itemize}
	%	\item (that they didn't find during stress testing)
	%\end{itemize}
	%%%% hopefully!
	\linegap
